---
title: "Gapminder Notebook"
output:
  html_document:
    df_print: paged
---

If needed, install the `kableExtra` package using:

```{r, eval = FALSE}
install.packages("kableExtra")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(kableExtra)
library(rmarkdown)
```

Make sure you have a `data` folder in your working directory.

If required, download the data by running the following in the console:

```
download.file(url="https://raw.githubusercontent.com/mvanrongen/2020_Gates_Reproducibility/master/data/processed/gapminder_1950_2018.csv", destfile="data/gapminder_1950_2018.csv"
```

## Loading the data
Change the location depending on where your file is located. Here the `gapminder_notebook.Rmd` file is located within a `scripts` folder, so we have to navigate up one level `(../)` to be able to access the `data` folder.

```{r load_data}
gapminder <- read_csv("../data/gapminder_1950_2018.csv")
```

### Dynamic tables using `rmarkdown`

```{r}
gapminder %>% 
  paged_table()
```

### Dynamic tables using `kableExtra`

```{r}
gapminder %>% 
  filter(world_region == "america",
         year == 2018) %>%
  select(country, life_expectancy) %>%
  arrange(desc(life_expectancy)) %>% 
  kable(caption = "Life expectancy in the Americas in 2018, descending") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE) %>% 
  scroll_box(width = "100%", height = "400px")
```












