---
title: "Gapminder Notebook"
output:
  html_document:
    df_print: paged
---

If needed, install the `kableExtra` package using:

```{r, eval = FALSE}
install.packages("kableExtra")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(kableExtra)
library(rmarkdown)
```

Make sure you have a `data` folder in your working directory.

If required, download the data by running the following in the console:

```
download.file(url="https://raw.githubusercontent.com/mvanrongen/2020_Gates_Reproducibility/master/data/processed/gapminder_1950_2018.csv", destfile="data/gapminder_1950_2018.csv"
```

## Loading the data

Change the location depending on where your file is located. Here the `gapminder_notebook.Rmd` file is located within a `scripts` folder, so we have to navigate up one level `(../)` to be able to access the `data` folder.

```{r load_data}
gapminder <- read_csv("../data/gapminder_1950_2018.csv")
```

### Dynamic tables using `rmarkdown`

```{r}
gapminder %>% 
  paged_table()
```

### Dynamic tables using `kableExtra`

```{r}
gapminder %>% 
  filter(world_region == "america",
         year == 2018) %>%
  select(country, life_expectancy) %>%
  arrange(desc(life_expectancy)) %>% 
  kable(caption = "Life expectancy in the Americas in 2018, descending") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE) %>% 
  scroll_box(width = "100%", height = "400px")
```

### Challenge 1

We are interested in the possible relationship between income and longevity. To investigate this, we can use the `income_per_person` and `life_expectancy` variables.

Use your `gapminder_notebook.Rmd` R Notebook file to explore this further. First, we are going to do some plotting to get an idea of what the data look like. When doing this, remember to do the following:

* use clear headings (`# Heading 1`, `## Heading 2` etc. to break up your code/explanations).
* name your ` ``` {r} ` chunks to help you navigate through your code

Try the following (preferably use pipes (`%>%` symbol) to get one data flow):

1. Filter the dataset for data from 2018 only (hint: use `filter`)
2. Plot using `ggplot()` a scatterplot of `income_per_person` vs `life_expectancy`
3. Colour the points by `income_group`
4. Facet the data by `world_region` (hint: use `facet_wrap`, see [documentation](https://ggplot2.tidyverse.org/reference/facet_wrap.html))

```{r, answer=TRUE, eval=TRUE}
gapminder %>% 
  filter(year == 2018) %>% 
  ggplot(aes(income_per_person, life_expectancy)) +
  geom_point(aes(colour = income_groups)) +
  facet_wrap(facets = vars(world_region))
```

Once you've managed to get this far, continue to Challenge 2.

### Challenge 2


```{r}
gapminder %>% 
  filter(year == 2018,
         world_region == "america",
         income_groups == "low_income")

gapminder %>% 
  filter(country_id == "hti") %>% 
  ggplot(aes(year, income_per_person)) +
  geom_point()
```











